---
title: "Penguin Flibber Lengths ANOVA"
author: "Nakoa Farrant"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages
```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)
```

Read in the penguins data
```{r, include = FALSE}
penguins <- read.csv("penguins.csv")
```

Our question: Is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago

three levels of species
continuous rand variable: flipper length

Maybe I can use one-way ANOVA...
- Independent observations
- Equal variance (Levene's test)
- Normality (histograms/qqplots)

first want to know how many observations were there in each group

```{r}
# with the penguins df, create a table that contains counts by species and sex
counts <- with(penguins, table(species, sex))

# data shows promise of being normally distributed given the counts. However, still need to do basic data analysis

penguins2 <- penguins %>% 
  # create a new column that has less elaborate species name
  mutate(
    common_name  = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% # remove species column
  filter(sex == "FEMALE")
```

