---
title: "Penguin Flibber Lengths ANOVA"
author: "Nakoa Farrant"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages
```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)
```

Read in the penguins data
```{r, include = FALSE}
penguins <- read.csv("penguins.csv")
```

Our question: Is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago

three levels of species
continuous rand variable: flipper length

Maybe I can use one-way ANOVA...
- Independent observations
- Equal variance (Levene's test)
- Normality (histograms/qqplots)

first want to know how many observations were there in each group

```{r}
# with the penguins df, create a table that contains counts by species and sex
counts <- with(penguins, table(species, sex))

# data shows promise of being normally distributed given the counts. However, still need to do basic data analysis

penguins2 <- penguins %>% 
  # create a new column that has less elaborate species name
  mutate(
    common_name  = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% # remove species column
  filter(sex == "FEMALE")
```

Exploratory graphs and Levene's test for equal variances

```{r}

hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name, scale = "free")

hists

qq_plots <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~ common_name, scale = "free")

qq_plots

# Based on qq-plots and histograms, observations for each group appear normally distributed. Also, n > 30 for all groups, so means will be normally distributed by CLT.
```
Evaluate variances and perform a Levene's test for equal variances

```{r}
# If the largest sample variances is < 4x greater than the smallest sample variance, then these are "close enough" to consider variances equal

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances # the largest variance (33.1) is not more than 4x greater than the smallest variance (15.2)

# do a formal hypothesis test for variance
# Levene's test (when you have more than two groups)
# leveneTest function in 'car' package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)
levene_penguin

# There is no significant difference in variances across species. If you have a very large sample size, you would get a significant difference in variances so it's a
# important to check if actually 4x bigger
```

One-way ANOVA for penguin flipper lengths by species (common_name)

# aov.test function in 'onewaytests' package DOES allow for inline references
# the native aov function does not
```{r}
# H0: Means across all groups are equal
# HA: At least two group means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

# reject Null hypothesis

# Post-hoc testing by Tukey's HSD
# H0: No significant difference in pairwise means for each pairing tested
penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

# Reject null hypothesis for all pair comparisons


```

Mean flipper lengths of female penguins differed significantly for all species studies (one-way anova, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with pot_hoc Tukey's HSD, $\alpha$ = 0.05).

